
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Step 1. Loading the relevant libraries
```{r libraries}
library(broom)
library(car)
library(tidyverse)
library(here)
library(ggplot2)
library(tidyr)
library(rstatix) # allows group_by()
library(effectsize) # a neat little package that provides the effect size
library(ggpubr) # allows us to examine QQ Plots 

```

# Step 2. Read in the data
```{r read in data}
demo_fb <- read_csv("erin_demo_fb_20240318.csv")
mw <- read_csv("mw_scores_20250223.csv")

demo_fb <- read_csv(here("mw2425_06data/erinRothwell-Wood/mw_2425_erw_processed", "erin_demo_fb_20240318.csv"))
mw <- read_csv(here("mw2425_06data/erinRothwell-Wood/mw_2425_erw_processed", "mw_scores_20250223.csv"))

```
```{r}
# Merge datasets on origID (participant ID)
merged_data <- mw %>%
  left_join(demo_fb, by = "origID")
```


################################## Removing the extreme outlier ####################################################

```{r}
library(dplyr)

data <- merged_data %>% filter(origID != 12538151)

```

################################# Gender split, mean age, SD age ######################################################

```{r}
# Going to use the demo_fb data set as the merged_data one has two rows for every participant
#First remove the extreme outlier participant from the demo_fb data set

demographics <- demo_fb %>% filter(origID != 12538151)

#Now find gender split, mean age, sd of age

# Count the number of men and women
table(demographics$gender)

# Calculate mean age
mean_age <- mean(demographics$age, na.rm = TRUE)
print(mean_age)

# Calculate standard deviation of age
sd_age <- sd(demographics$age, na.rm = TRUE)
print(sd_age)
```

################################ Histogram to represent the skewed age range ###################################
```{r}
ggplot(demographics, aes(x = age, fill = gender)) +
  geom_histogram(binwidth = 2.5, position = "stack", color = "black") +
  labs(title = "Age Distribution by Gender", x = "Age", y = "Count") +
  scale_fill_manual(values = c("blue", "pink")) +  # Adjust colors as needed
  theme_minimal()
```

############################## Calculating the means and sds for all relevant variables ##################################################

```{r}

# Group by condition and calculate the mean and standard deviation for all relevant variables
stats_by_condition <- data %>%
  group_by(condition) %>%
  summarise(
    mean_mm_acc = mean(mm_acc, na.rm = TRUE),
    sd_mm_acc = sd(mm_acc, na.rm = TRUE),
    mean_mm_rt = mean(mm_rt, na.rm = TRUE),
    sd_mm_rt = sd(mm_rt, na.rm = TRUE),
    mean_mw_freq = mean(mw_freq, na.rm = TRUE),
    sd_mw_freq = sd(mw_freq, na.rm = TRUE),
    mean_mw_intent_freq = mean(mw_intent_freq, na.rm = TRUE),
    sd_mw_intent_freq = sd(mw_intent_freq, na.rm = TRUE)
  )

# Print result
print(stats_by_condition)


```

############################### Creating Bar Charts to represent the means/sd's ##########################################

```{r}

# Compute summary statistics for mm_acc
mm_acc_stats <- data %>%
  group_by(condition) %>%
  summarise(
    mean_mm_acc = mean(mm_acc, na.rm = TRUE),
    sd_mm_acc = sd(mm_acc, na.rm = TRUE)
  )

# Create bar plot with error bars
ggplot(mm_acc_stats, aes(x = condition, y = mean_mm_acc, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +  # Bar plot
  geom_errorbar(aes(ymin = mean_mm_acc - sd_mm_acc, ymax = mean_mm_acc + sd_mm_acc),
                width = 0.2, position = position_dodge(0.9)) +  # Error bars for SD
  labs(title = "Mean mm_acc by Condition",
       x = "Condition",
       y = "Mean Accuracy (mm_acc)") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove legend since we only have 2 conditions

# Compute summary statistics for mm_rt
mm_rt_stats <- data %>%
  group_by(condition) %>%
  summarise(
    mean_mm_rt = mean(mm_rt, na.rm = TRUE),
    sd_mm_rt = sd(mm_rt, na.rm = TRUE)
  )

# Create bar plot with error bars
ggplot(mm_rt_stats, aes(x = condition, y = mean_mm_rt, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +  # Bar plot
  geom_errorbar(aes(ymin = mean_mm_rt - sd_mm_rt, ymax = mean_mm_rt + sd_mm_rt),
                width = 0.2, position = position_dodge(0.9)) +  # Error bars for SD
  labs(title = "Mean mm_rt by Condition",
       x = "Condition",
       y = "Mean Reaction Time (mm_rt)") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove legend since we only have 2 conditions


# Compute summary statistics for mw_freq
mw_freq_stats <- data %>%
  group_by(condition) %>%
  summarise(
    mean_mw_freq = mean(mw_freq, na.rm = TRUE),
    sd_mw_freq = sd(mw_freq, na.rm = TRUE)
  )

# Create bar plot with error bars
ggplot(mw_freq_stats, aes(x = condition, y = mean_mw_freq, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +  # Bar plot
  geom_errorbar(aes(ymin = mean_mw_freq - sd_mw_freq, ymax = mean_mw_freq + sd_mw_freq),
                width = 0.2, position = position_dodge(0.9)) +  # Error bars for SD
  labs(title = "Mean mw_freq by Condition",
       x = "Condition",
       y = "Mean Mind-Wandering Frequency (mw_freq)") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove legend since we only have 2 conditions



# Compute summary statistics for mw_intent_freq
mw_intent_freq_stats <- data %>%
  group_by(condition) %>%
  summarise(
    mean_mw_intent_freq = mean(mw_intent_freq, na.rm = TRUE),
    sd_mw_intent_freq = sd(mw_intent_freq, na.rm = TRUE)
  )

# Create bar plot with error bars
ggplot(mw_intent_freq_stats, aes(x = condition, y = mean_mw_intent_freq, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +  # Bar plot
  geom_errorbar(aes(ymin = mean_mw_intent_freq - sd_mw_intent_freq, ymax = mean_mw_intent_freq + sd_mw_intent_freq),
                width = 0.2, position = position_dodge(0.9)) +  # Error bars for SD
  labs(title = "Mean mw_intent_freq by Condition",
       x = "Condition",
       y = "Mean Intentional Mind-Wandering Frequency (mw_intent_freq)") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove legend since we only have 2 conditions


#Reaction Time Box Plot with individual points
ggplot(data, aes(x = condition, y = mm_rt, color = condition)) +
  geom_jitter(alpha = 0.5, width = 0.2) +  # Adds jittered points
  geom_boxplot(alpha = 0.2, outlier.shape = NA) +  # Adds boxplot without outliers
  labs(title = "Scatter and Boxplot of mm_rt by Condition",
       x = "Condition",
       y = "Reaction Time (mm_rt)") +
  theme_minimal()


```
######## mm_acc by condition t-test#################
```{r}
t.test(mm_acc ~ condition, paired = TRUE, data = data)
```













####################################### Research Question 1 ##########################################################
How is MW freq affected by easy + diff conditions during mental maths qs




```{r}
# Calculate mean difference (Difficult - Easy)
mw_freq_stats$mean_mw_freq[mw_freq_stats$condition == "Difficult"] - 
mw_freq_stats$mean_mw_freq[mw_freq_stats$condition == "Easy"]


t.test(mw_freq ~ condition, paired = TRUE, data = data)


# Compute mean difference
mean_diff <- mean(data$mw_freq[data$condition == "Difficult"] - 
                  data$mw_freq[data$condition == "Easy"])

# Compute standard deviation of differences
sd_diff <- sd(data$mw_freq[data$condition == "Difficult"] - 
              data$mw_freq[data$condition == "Easy"])

# Calculate Cohen’s d
cohen_d <- mean_diff / sd_diff
cohen_d



```

A paired samples t-test was conducted to compare mind-wandering frequency during difficult and easy questions. The results indicated that participants mind-wandered significantly more during difficult questions (M = 36.92%, SD = 29.50%) than during easy questions (M = 23.26%, SD = 24.18%), t(41) = 3.41, p < .001, d = 0.53. The mean difference was 12.80 (95% CI [5.22, 20.37]), suggesting a moderate increase in mind-wandering frequency when participants were answering more difficult questions.

############ Visualise results of paired t-test ########################

```{r}

library(ggplot2)

# Compute summary statistics (mean and standard error)
summary_data <- data %>%
  group_by(condition) %>%
  summarise(mean_mw = mean(mw_freq), 
            se_mw = sd(mw_freq) / sqrt(n()))  # Standard error

# Bar plot
ggplot(summary_data, aes(x = condition, y = mean_mw, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = mean_mw - se_mw, ymax = mean_mw + se_mw), 
                width = 0.2, color = "black") +  # Error bars
  theme_minimal() +
  labs(title = "Mind-Wandering Frequency by Question Difficulty",
       x = "Condition",
       y = "Mean Mind-Wandering Frequency") +
  scale_fill_manual(values = c("blue", "red")) +
  theme(legend.position = "none")



ggplot(data, aes(x = condition, y = mw_freq, fill = condition)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +  # Boxplot without outlier dots
  geom_jitter(aes(color = condition), width = 0.2, alpha = 0.6) +  # Individual points
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, color = "black") +  # Mean point
  theme_minimal() +
  labs(title = "Mind-Wandering Frequency by Question Difficulty",
       x = "Condition",
       y = "Mind-Wandering Frequency") +
  scale_fill_manual(values = c("blue", "red")) +
  theme(legend.position = "none")


```




########################### Checking assumptions of normality #############################################

```{r}
diff_mw <- data$mw_freq[data$condition == "Difficult"] - 
           data$mw_freq[data$condition == "Easy"]


hist(diff_mw, main = "Histogram of Differences", xlab = "Difference (Difficult - Easy)", col = "lightblue", breaks = 10)


shapiro.test(diff_mw)


```

A Shapiro-Wilk test indicated that the assumption of normality was violated (W = 0.88, p < .001). Therefore, a Wilcoxon Signed-Rank Test was conducted instead of a paired samples t-test.

```{r}
wilcox.test(mw_freq ~ condition, paired = TRUE, data = data)

```
A Wilcoxon Signed-Rank Test indicated a significant difference in mind-wandering frequency between Easy and Difficult questions, V = 449.5, p = .002. These results suggest that participants mind-wandered more during Difficult questions compared to Easy ones.

####################Multiple regression#################################

```{r}
RQ_1_Reg <- lm(mw_freq ~ condition + set.x + age + gender, data = data)
# View regression results
summary(model)
```

1. Effect of Condition
Participants in the "Easy" condition had 12.80% lower mind-wandering frequency (mw_freq) compared to those in the "Difficult" condition (p = 0.0257).
This suggests that harder tasks lead to more mind-wandering, which aligns with cognitive load theories—easier tasks may allow participants to stay more engaged.

2. Effect of Set.x
Set (1 vs. 2) had no significant effect (p = 0.4159), meaning Set 1 and Set 2 were equally difficult/easy as intended

3. Effect of Age
Age did not significantly predict mind-wandering frequency (p = 0.3396).
This suggests that younger and older participants mind-wander at similar rates in this task.

4. Effect of Gender
Women had 15.20% higher mind-wandering frequency compared to men (p = 0.0274).
This is an interesting finding! It could be due to differences in cognitive processing, engagement.















################################################### Research Question 2 #################################################################
Is there a difference in the type of MW in the easy vs difficult conditions


```{r}
library(tidyr)
library(dplyr)

# Reshape data to wide format
wide_data <- data %>%
  select(origID, condition, mw_intent_freq) %>%
  pivot_wider(names_from = condition, values_from = mw_intent_freq)

# Remove rows with missing values (paired t-test requires complete cases)
wide_data <- na.omit(wide_data)

t.test(wide_data$Easy, wide_data$Difficult, paired = TRUE)


```

A paired-samples t-test was conducted to compare intentional mind-wandering frequency in the Easy and Difficult conditions. The results indicated no significant difference, t(26) = -0.77, p = .447, d = -0.15, 95% CI [-26.25, 11.91]. On average, participants reported slightly lower intentional mind-wandering in the Difficult condition (M = , SD = ) compared to the Easy condition (M = , SD = ), but this difference was not statistically significant.

```{r}
shapiro.test(wide_data$Easy - wide_data$Difficult)

```

t-test is valid 


```{r}

# Calculate the difference between conditions
differences <- wide_data$Easy - wide_data$Difficult

# Compute mean difference
mean_diff <- mean(differences)

# Compute standard deviation of the differences
sd_diff <- sd(differences)

# Compute Cohen's d
cohens_d <- mean_diff / sd_diff

# Print result
print(cohens_d)

```

There is no significant difference in mw_intent_freq between the Easy and Difficult conditions (p = 0.45).
The effect size (d = -0.15) confirms that the difference is very tiny.

############# visualising data ###########
```{r}
library(ggplot2)

ggplot(data, aes(x = condition, y = mw_intent_freq)) +
  geom_boxplot(aes(fill = condition), alpha = 0.5) +  
  geom_jitter(width = 0.2, alpha = 0.5) +  
  theme_minimal() +
  labs(title = "Intentional Mind-Wandering Frequency by Condition",
       x = "Condition", y = "Intentional Mind-Wandering Frequency")

```

################### Multiple Regression ####################################
```{r}
long_data <- data %>%
  select(origID, condition, set.x, age, gender, mw_intent_freq)  # Keep relevant variables

model_intent <- lm(mw_intent_freq ~ condition + set.x + age + gender, data = long_data)
summary(model_intent)

```

None of the predictors are significant
Condition (Easy vs. Difficult): Participants in the Easy condition had 13.05% lower intentional mind-wandering than those in the Difficult condition, but this difference is not statistically significant (p = 0.196).
Set.x: No significant effect (p = 0.481), confirming that Set 1 and Set 2 were equally difficult/easy as expected. 
Age: Small positive effect (β = 0.4452) but not significant (p = 0.177) - older participants might intentionally mind-wander more, but this is unclear.
Gender (Women vs. Men): Women reported 11.73% lower intentional mind-wandering than men, but this effect is not significant (p = 0.350).

R² = 0.0706 - Only 7.06% of the variance in intentional mind-wandering is explained by this model.
Adjusted R² = 0.0076 - After adjusting for the number of predictors, the model explains almost no variance.
F-statistic = 1.121, p = 0.3555 - The overall model is NOT statistically significant, meaning that these predictors do not explain intentional mind-wandering well.

```{r}
plot(model_intent)

```


Unlike previous model predicting mw_freq, where condition and gender were significant, this model suggests that intentional mind-wandering is not strongly related to these variables.


Task difficulty may affect men and women differently???
```{r}
model_interaction <- lm(mw_intent_freq ~ condition * gender + set.x + age, data = long_data)
summary(model_interaction)

```

Main Effects (No Significant Predictors)
Condition (Easy vs. Difficult) - Participants in the Easy condition reported 33.53% lower intentional mind-wandering than those in the Difficult condition, but this effect is not significant (p = 0.1395).
Gender (Women vs. Men) - Women reported 24.88% lower intentional mind-wandering than men, but this effect is also not significant (p = 0.1700).

Age & Set.x - No significant effects (p > 0.1).
Interaction Effect (Condition × Gender)
The interaction term conditionEasy:genderWoman = 25.56 (p = 0.3108)
This suggests that the gender difference in intentional mind-wandering depends on task difficulty, but the effect is not statistically significant.

R² = 0.087 - Only 8.7% of variance in intentional mind-wandering is explained.
Adjusted R² = 0.008 - After adjusting for predictors, the model explains almost nothing.
F-statistic (1.106, p = 0.3671) - The overall model is not statistically significant.

No strong evidence that task difficulty (condition) or gender affects intentional mind-wandering.
No evidence of an interaction between condition and gender.
Other factors likely explain intentional mind-wandering better (e.g., motivation, cognitive control, fatigue).

```{r}
plot(model_interaction$fitted.values, model_interaction$residuals)
abline(h = 0, col = "red")


plot(model_interaction, which = 3)  # Scale-Location plot


qqnorm(model_interaction$residuals)
qqline(model_interaction$residuals, col = "red")  # Reference line


```

Linearity - Is good
Homoscedasticity - is good
Normality of residuals - not good - but deviation is moderate
So There is a mild violation 











################################################# Research Question 3 ###########################################################
Test Performance in relation to Mind Wandering


#################### Correlation: to test mind-wandering against test performance#######################

```{r}
cor_test <- cor.test(data$mw_freq, data$mm_acc, method = "pearson")
print(cor_test)

```
Correlation coefficient (r) = -0.5088

This indicates a moderate negative correlation between mw_freq and mm_acc 
As mind-wandering increases, test performance decreases.
p-value = 7.748e-07 

This is extremely significant (p < 0.05), meaning the relationship between mw_freq and mm_acc is unlikely to be due to chance.
95% Confidence Interval: [-0.6521, -0.3304]

We are 95% confident that the true correlation is between -0.652 and -0.330, further confirming a negative relationship.

There is strong statistical evidence that higher mind-wandering leads to lower test performance.
However, correlation does not imply causation. Other factors could also play a role.

####### visulaising the results ###################

```{r}

# Compute correlation coefficient
cor_value <- cor(data$mw_freq, data$mm_acc)

# Create scatter plot with regression line
ggplot(data, aes(x = mw_freq, y = mm_acc)) +
  geom_point(color = "blue", alpha = 0.6) +  # Scatter points
  geom_smooth(method = "lm", color = "red", se = TRUE) +  # Regression line with confidence interval
  annotate("text", x = max(data$mw_freq), y = min(data$mm_acc), 
           label = paste("r =", round(cor_value, 3)), 
           color = "black", hjust = 1, size = 5) +  # Display correlation coefficient
  labs(title = "Correlation between MW Frequency and MM Accuracy",
       x = "MW Frequency (%)",
       y = "MM Accuracy (%)") +
  theme_minimal()

```



```{r}


# Run multiple regression model
model <- lm(mm_acc ~ mw_freq + condition + set.x + age + gender, data = data)

# View the results
summary(model)

```

Only mw_freq is statistically significant (p = 1.54e-05).
This strongly suggests higher mind-wandering negatively impacts test performance.
Condition, set, age, and gender are not significant (p > 0.05), meaning they do not have a strong independent effect on test scores in this dataset.
Multiple R² = 0.3084
The model explains 30.84% of the variance in test performance.
This suggests that other factors (not included in this model) also play a role.
Adjusted R² = 0.2641
Adjusted for the number of predictors, the model still explains 26.41% of the variance.

###########################Checking Assumptions#########################
```{r}

# Scatterplots for numeric predictors
ggplot(data, aes(x = mw_freq, y = mm_acc)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: MW Frequency vs MM Accuracy")

ggplot(data, aes(x = age, y = mm_acc)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: Age vs MM Accuracy")

```


```{r}
plot(model$fitted.values, resid(model),
     main = "Residual Plot: Checking Homoscedasticity",
     xlab = "Fitted Values", ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)

```
```{r}
qqnorm(resid(model))
qqline(resid(model), col = "red")

shapiro.test(resid(model))

```

The residuals are NOT normally distributed - violates the normality assumption
This could impact the reliability of hypothesis tests and confidence intervals in regression model.





```{r}
# Run multiple regression with interaction
model_interaction <- lm(mm_acc ~ mw_freq * condition + set.x + age + gender, data = data)

# View summary results
summary(model_interaction)


```
note to self: the write up of this is in results word doc in the my files

################################ Checking Assumptions ############################################
```{r}
# Check linearity for mw_freq
ggplot(data, aes(x = mw_freq, y = mm_acc, color = condition)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: MW Frequency vs MM Accuracy")

# Check linearity for age
ggplot(data, aes(x = age, y = mm_acc)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: Age vs MM Accuracy")


plot(model_interaction$fitted.values, resid(model_interaction),
     main = "Residual Plot: Checking Homoscedasticity",
     xlab = "Fitted Values", ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)

hist(resid(model_interaction), main = "Histogram of Residuals", col = "lightblue")

qqnorm(resid(model_interaction))
qqline(resid(model_interaction), col = "red")

shapiro.test(resid(model_interaction))


```
The residuals are NOT normally distributed - violates the normality assumption
This could impact the reliability of hypothesis tests and confidence intervals in regression model.





############# Intentionality of Mind-wandering against task performance ####################################
```{r}
# Run multiple regression with additional predictors
model_intent_full <- lm(mm_acc ~ mw_intent_freq + condition + set.x + age + gender, data = data)

# View summary results
summary(model_intent_full)

```

Itentional mind-wandering (mw_intent_freq) does not significantly impact test performance (p =0.276)
ndition (Easy vs. Difficult) is a significant predictor (p=0.024), with better test performance in the Easy condition.
Set, age, and gender do not significantly impact test scores.

#############################Checking Assumptions###################################
```{r}
library(ggplot2)

# Check linearity for mw_intent_freq
ggplot(data, aes(x = mw_intent_freq, y = mm_acc)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: MW Intent Frequency vs MM Accuracy")

# Check linearity for age
ggplot(data, aes(x = age, y = mm_acc)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: Age vs MM Accuracy")


plot(model_intent_full$fitted.values, resid(model_intent_full),
     main = "Residual Plot: Checking Homoscedasticity",
     xlab = "Fitted Values", ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)


hist(resid(model_intent_full), main = "Histogram of Residuals", col = "lightblue")

qqnorm(resid(model_intent_full))
qqline(resid(model_intent_full), col = "red")

shapiro.test(resid(model_intent_full))

```



```{r}

# Run regression with interaction term
model_intent_interaction <- lm(mm_acc ~ mw_intent_freq * condition + set.x + age + gender, data = data)

# View summary results
summary(model_intent_interaction)

```
note to self: the write up of this is in results word doc in the my files

########################## Checking Assumptions ######################################
```{r}
library(ggplot2)

# Check linearity for mw_intent_freq
ggplot(data, aes(x = mw_intent_freq, y = mm_acc, color = condition)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: MW Intent Frequency vs MM Accuracy")

# Check linearity for age
ggplot(data, aes(x = age, y = mm_acc)) + 
  geom_point() + 
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Linearity Check: Age vs MM Accuracy")


plot(model_intent_interaction$fitted.values, resid(model_intent_interaction),
     main = "Residual Plot: Checking Homoscedasticity",
     xlab = "Fitted Values", ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)


hist(resid(model_intent_interaction), main = "Histogram of Residuals", col = "lightblue")

qqnorm(resid(model_intent_interaction))
qqline(resid(model_intent_interaction), col = "red")

shapiro.test(resid(model_intent_interaction))

```

Residuals do not follow a normal distribution, violating the normality assumption of multiple regression. This can affect the reliability of hypothesis and confidence intervals.
