---
title: "Analysis for mind wandering during mental maths task"
author: "Erin and Margriet"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab activity 2 - Statistics anxiety and engagement in module activities

# Step 1. Loading the relevant libraries
```{r libraries}
library(broom)
library(car)
library(tidyverse)
library(here)
library(rstatix)
```

# Step 2. Read in the data
```{r read in data}
demo_fb <- read_csv(here("mw2425_06data/erinRothwell-Wood/mw_2425_erw_processed", "erin_demo_fb_20240318.csv"))
mw_scores <- read_csv(here("mw2425_06data/erinRothwell-Wood/mw_2425_erw_processed", "mw_scores_20250223.csv"))

head(mw_scores) # Look at the data frames
head(demo_fb)
```

# Step 3: Calculating descriptive statistics

```{r descriptives}
descriptives_mm <- mw_scores %>%
  filter(origID != 12538151) %>%
  group_by(condition, set) %>%
  summarise(mean_mm_acc = mean(mm_acc, na.rm = TRUE),
            sd_mm_acc = sd(mm_acc, na.rm = TRUE),
            min_mm_acc = min(mm_acc, na.rm = TRUE),
            max_mm_acc = max(mm_acc, na.rm = TRUE),
            mean_mm_rt = mean(mm_rt, na.rm = TRUE),
            sd_mm_rt = sd(mm_rt, na.rm = TRUE),
            min_mm_rt = min(mm_rt, na.rm = TRUE),
            max_mm_rt = max(mm_rt, na.rm = TRUE),
            mean_mw_freq = mean(mw_freq, na.rm = TRUE),
            sd_mw_freq = sd(mw_freq, na.rm = TRUE),
            min_mw_freq = min(mw_freq, na.rm = TRUE),
            max_mw_freq = max(mw_freq, na.rm = TRUE),
            mean_mw_intent_freq = mean(mw_intent_freq, na.rm = TRUE),
            sd_mw_intent_freq = sd(mw_intent_freq, na.rm = TRUE),
            min_mw_intent_freq = min(mw_intent_freq, na.rm = TRUE),
            max_mw_intent_freq = max(mw_intent_freq, na.rm = TRUE))
descriptives_mm

```

# Step 4: Visualise the data

```{r boxplots}
ggplot(data = mw_scores, mapping = aes(x = mm_rt)) +
  geom_histogram()

mw_scores_excl <- mw_scores %>%
  filter(origID != 12538151) %>%
  filter(mm_rt < 20000) %>%
  mutate(origID = as.factor(origID),
         condition = as.factor(condition),
         set = as.factor(set))

ggplot(data = mw_scores_excl, mapping = aes(x = mm_rt)) +
  geom_histogram()

mw_scores_excl$condition <- factor(mw_scores_excl$condition, levels = c('Easy', 'Difficult'))

ggplot(data = mw_scores_excl, mapping = aes(x = condition, y = mm_acc)) +
  geom_boxplot() +
  labs(x = "Condition", y = "Accuracy (%)") +
  theme_bw()

ggplot(data = mw_scores_excl, mapping = aes(x = condition, y = mm_rt)) +
  geom_boxplot() +
  labs(x = "Condition", y = "Reaction time (ms)") +
  theme_bw()

ggplot(data = mw_scores_excl, mapping = aes(x = condition, y = mw_freq)) +
  geom_boxplot() +
  labs(x = "Condition", y = "Mind wandering frequency (%)") +
  theme_bw()

ggplot(data = mw_scores_excl, mapping = aes(x = condition, y = mw_intent_freq)) +
  geom_boxplot() +
  labs(x = "Condition", y = "Frequency of intent (%)") +
  theme_bw()

```

# Step 5: Build the regression model ----------------------------------

```{r model}
# Test whether task performance (accuracy) is different on easy and difficult blocks - paired t-test with mm_acc
mod_mm_acc <- lm(mm_acc ~ condition + set, data = mw_scores_excl)
mod_mm_acc_summary <- summary(mod_mm_acc)
mod_mm_acc_summary

t_mm_acc <- mw_scores_excl %>%
  t_test(mm_acc ~ condition, detailed = TRUE)
t_mm_acc

# add two more t-tests for the other two comparisions
# Test whether task performance (reaction times) is different on easy and difficult blocks - paired t-test with mm_rt
mod_mm_rt <- lm(mm_rt ~ condition + set, data = mw_scores_excl)
mod_mm_rt_summary <- summary(mod_mm_rt)
mod_mm_rt_summary

t_mm_rt <- mw_scores_excl %>%
  t_test(mm_rt ~ condition, detailed = TRUE)
t_mm_rt 


# Test whether mind wandering frequency is different during easy and difficult blocks - multiple regression with mw_freq as DV and condition as the IV (and set as IV)

mod_mw <- lm(mw_freq ~ condition + set + age + condition*age, data = mw_scores_excl)
mod_mw_summary <- summary(mod_mw)
mod_mw_summary

# Test whether mind wandering intention is different during easy and difficult blocks - multiple regression with mw_intent as DV and condition as the IV (and set as IV)

mod_intent <- lm(mw_intent_freq ~ condition + set, data = mw_scores_excl)
mod_intent_summary <- summary(mod_intent)
mod_intent_summary

cor_mwfreq <- cor.test(mw_scores_excl$mw_freq, 
                    mw_scores_excl$mm_acc, 
                    method = "pearson", 
                    alternative = "two.sided") %>% 
  tidy()
cor_mwfreq

ggplot(data = mw_scores_excl, mapping = aes(x = mm_acc, y = mw_freq, colour = condition)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Accuracy (%)", y = "Mind wandering frequency (%)") +
  theme_bw()

ggplot(data = mw_scores_excl, mapping = aes(x = mm_rt, y = mw_freq, colour = condition)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Reaction time (ms)", y = "Mind wandering frequency (%)") +
  theme_bw()
```

# Step 6: Checking assumptions

```{r assumptions}
crPlots(mod)                  # Plot linear line and line that best fits the data to check the relationship between outcome and predictor is linear

qqPlot(mod$residuals)         # Create qq-plot to check residuals are normally distributed

residualPlot(mod)             # Create residual plot to check residual show homoscedasticity
```

# Step 7: Write up

**A simple linear regression was performed with engagement (*M* = 4.54, *SD* = 0.56) as the outcome variable and statistics anxiety (*M* = 2.08, *SD* = 0.56) as the predictor variable. The results of the regression indicated that the model significantly predicted course engagement (*F*(1, 35) = 11.99, *p* \< .001, *R\^2* = 0.25), accounting for 25% of the variance. Anxiety was a significant negative predictor (*beta* = -2.17, *p* \< 0.001): as anxiety increased, course engagement decreased.**