---
title: "Analysis for mind wandering during mental maths task"
author: "Erin and Margriet"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab activity 2 - Statistics anxiety and engagement in module activities

# Step 1. Loading the relevant libraries
```{r libraries}
library(broom)
library(car)
library(tidyverse)
library(here)
```

# Step 2. Read in the data
```{r read in data}
demo_fb <- read_csv("erin_demo_fb_20240318.csv")
mw <- read_csv("mw_scores_20250215.csv")

head(stars) # Look at the data frames
head(engage)
```

# Step 3: Calculating descriptive statistics

```{r descriptives}
descriptives_demo <- demo_fb %>%
  summarise(mean_anx = mean(mean_anxiety, na.rm = TRUE),
            sd_anx = sd(mean_anxiety, na.rm = TRUE),
            mean_weeks = mean(n_weeks, na.rm = TRUE),
            sd_weeks = sd(n_weeks, na.rm = TRUE))
descriptives_demo


descriptives_mw

group_by() #condition
```

# Step 4: Visualise the data

```{r scatterplot}
ggplot(joined, aes(x = mean_anxiety, y = n_weeks)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Anxiety", y = "Engagement") +
  theme_bw()
```

# Step 5: Build the regression model ----------------------------------

```{r model}
# Test whether task performance (accuracy) is different on easy and difficult blocks - paired t-test with mm_acc

# Test whether task performance (reaction times) is different on easy and difficult blocks - paired t-test with mm_rt

# Test whether mind wandering frequency is different during easy and difficult blocks - multiple regression with mw_freq as DV and condition as the IV (and set as IV)

mod <- lm(mw_freq ~ condition + set, data = mw)
mod_summary <- summary(mod)
mod_summary

# Test whether mind wandering intention is different during easy and difficult blocks - multiple regression with mw_intent as DV and condition as the IV (and set as IV)

mod <- lm(mw_freq ~ condition + set, data = mw)
mod_summary <- summary(mod)
mod_summary
```

# Step 6: Checking assumptions

```{r assumptions}
crPlots(mod)                  # Plot linear line and line that best fits the data to check the relationship between outcome and predictor is linear

qqPlot(mod$residuals)         # Create qq-plot to check residuals are normally distributed

residualPlot(mod)             # Create residual plot to check residual show homoscedasticity
```

# Step 7: Write up

**A simple linear regression was performed with engagement (*M* = 4.54, *SD* = 0.56) as the outcome variable and statistics anxiety (*M* = 2.08, *SD* = 0.56) as the predictor variable. The results of the regression indicated that the model significantly predicted course engagement (*F*(1, 35) = 11.99, *p* \< .001, *R\^2* = 0.25), accounting for 25% of the variance. Anxiety was a significant negative predictor (*beta* = -2.17, *p* \< 0.001): as anxiety increased, course engagement decreased.**