---
title: "PSYC400 Analysis Script"
author: "Molly Pugh"
date: "2024-07-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Clear environment

```{r}
rm(list=ls())
```

# Load libraries

```{r}
library(broom)
library(effects)
library(gridExtra)
library(here)
library(lattice)
library(knitr)
library(lme4)
library(MuMIn)
library(sjPlot)
library(tidyverse)
library(memisc)
library(ggplot2)
library(Rmisc)
```

# Read in the data.

```{r}
DOCS <- read_csv("DOCS_Preview_Data.csv")
MWT <- read_csv("Mind_Wandering_Task_Preview_Data.csv")
DQ <- read_csv("Demographics_Preview_data.csv")
```
# Tidy the data.

## Demographics Questionnaire

### Select columns required from MWT dataset

```{r}
DQ_2 <- DQ %>%
  dplyr::select("Participant Private ID", "Response",  "Question Key")
```

### Remove nas

```{r}
DQ_3 <- DQ_2 %>%
  na.omit()
```

### Change into wide format

```{r}
DQ_3_wider <- DQ_3 %>%
  pivot_wider(names_from = "Question Key",
              values_from = "Response")
```

### Select final columns required

```{r}
DQ_4 <- DQ_3_wider %>%
  dplyr::select("Participant Private ID", "Age",  "Gender", "multilingual", "OCD_Diagnosis")
```

## DOCS Questionnaire

### Select columns required from DOCS dataset

```{r}
DOCS_2 <- DOCS %>%
  dplyr::select("Participant Private ID", "Key", "Response", "Response Type")
```

### Remove nas

```{r}
DOCS_3 <- DOCS_2 %>%
  na.omit()
```

### Filter Key Quantised and Response Type

```{r}
DOCS_4 <- DOCS_3 %>%
  filter(Key == "quantised", `Response Type` == "response")
```

### Select needed columns

```{r}
DOCS_5 <- DOCS_4 %>%
  dplyr::select("Participant Private ID","Response")
```

### Change "Response" to a numeric variable

```{r}
DOCS_5$Response <- as.numeric(DOCS_5$Response)
```

### Create total DOCS score for each participant

```{r}
DOCS_5$DOCS_Score <- DOCS_5 %>%
  group_by("Participant Private ID") %>%
  summarise(sum(Response))
```

## Mind Wandering Task

### Select columns required from MWT dataset

```{r}
MWT_2 <- MWT %>%
  dplyr::select("Participant Private ID", "Response",  "Spreadsheet: condition", "Spreadsheet: passage", "Spreadsheet: line", "Spreadsheet: set", "Spreadsheet: correct_response", "Object Name")
```

### Filter Responses

```{r}
MWT_3 <- MWT_2 %>%
  filter(Response == c("Intentional", "On Task", "Unintentional"), 
         `Response Type` == "response")
```

# Centre and Standardise Continuous Variables

```{r}
dat_4_z <- dat_4 %>%
  mutate(zconf = (conf - mean(conf, na.rm = TRUE)) / sd(conf))
```

# Demographics

### Initial Gender count

```{r}
initial_gender <- DQ_4 %>%
  count("Gender")
```

### Initial Mean Age

#### Change Age to Numeric

```{r}
DQ_4$Age <- as.numeric(DQ_4$Age)
```

```{r}
initial_mean_age <-
  mean(DQ_4$Age)
```

### Initial SD of Age

```{r}
initial_sd <-
  sd(DQ_4$Age)
```

### Initial Age Range

```{r}
initial_range <- range(DQ_4$Age)
```

### Demographic with exclusions

```{r}
DQ_5
```

### Final Gender Count

```{r}
gender <- DQ_5 %>%
  count("Gender")
```

### Final Mean age 

```{r}
mean_age <-
  mean(DQ_5$Age)
```

### Final SD Age

```{r}
Age_sd <-
  sd(DQ_5$Age)
```

### Final Age Range

```{r}
Age_range <- range(DQ_5$Age)
```

# Affect of OCD and Task Difficulty on mind wandering overall

## Finding the most appropriate model

### Random effects of participants and stimulus items on intercepts

```{r}
model_1.1 <- glmer(MW_overall ~ DOCS_Score + 
                     
                     Task_Difficulty + DOCS_Score:Task_Difficulty +
                               
                               (1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_1.1)
```

### Random effects of participants on intercepts and slope of DOCS_Score, and stimulus items on intercepts

```{r}
model_2.1 <- glmer(MW_overall ~ DOCS_Score + 
                     
                     Task_Difficulty + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_2.1)
```

### Comparing model_1.1 to model_2.1

```{r}
anova(model_1.1, model_2.1)
```

### Random effects of participants on intercepts and slope of DOCS_Score, and stimulus items on intercepts with covariances

```{r}
model_3.1 <- glmer(MW_overall ~ DOCS_Score + 
                     
                     Task_Difficulty + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 1 | ResponseId) + 
                               
                               (1 | Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_3.1)
```

### Comparing model_3.1 to model_4.1

```{r}
anova(model_2.1, model_3.1)
```

## Visualising the results

# Affect of OCD and task difficulty on mind wandering intention

## Finding the most appropriate model

### Random effects of participants and stimulus items on intercepts

```{r}
model_1.2 <- glmer(MW_intention ~ DOCS_Score + 
                     
                     Task_Difficulty + DOCS_Score:Task_Difficulty +
                               
                               (1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_1.2)
```

### Random effects of participants on intercepts and slope of DOCS_Score, and stimulus items on intercepts

```{r}
model_2.2 <- glmer(MW_intention ~ DOCS_Score + 
                     
                     Task_Difficulty + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_2.2)
```

### Comparing model_1.1 to model_2.1

```{r}
anova(model_1.2, model_2.2)
```

### Random effects of participants on intercepts and slope of DOCS_Score, and stimulus items on intercepts with covariances

```{r}
model_3.2 <- glmer(MW_intention ~ DOCS_Score + 
                     
                     Task_Difficulty + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 1 | ResponseId) + 
                               
                               (1 | Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_3.2)
```

### Comparing model_2.2 to model_3.2

```{r}
anova(model_2.2, model_3.2)
```

## Visualising the results

# Affect of OCD, task difficulty and mind wandering on Reading Comprehension

## Finding the most appropriate model

### Random effects of participants and stimulus items on intercepts

```{r}
model_1.3 <- glmer(Acc ~ DOCS_Score + 
                     
                     Task_Difficulty + MW + DOCS_Score:Task_Difficulty +
                               
                               (1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_1.3)
```


### Random effects of participants on intercepts and task difficulty slope, and stimulus items on intercepts

```{r}
model_2.3 <- glmer(Acc ~ DOCS_Score + 
                     
                     Task_Difficulty + MW + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty)+ 1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_2.3)
```

### Comparing model_1.3 to model_2.3

```{r}
anova(model_1.3, model_2.3)
```

### Random effects of participants on intercepts and task difficulty and mind wandering slope, and stimulus items on intercepts

```{r}
model_3.3 <- glmer(Acc ~ DOCS_Score + 
                     
                     Task_Difficulty + MW + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 
                                  dummy(MW) + 1 || ResponseId) + 
                               
                               (1 || Stimulus),
                             
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_3.3)
```

### Comparing model_2.3 to model_3.3

```{r}
anova(model_2.3, model_3.3)
```

### Random effects of participants on intercepts and task difficulty and mind wandering slope, and stimulus items on intercepts and mind wandering slope

```{r}
model_4.3 <- glmer(Acc ~ DOCS_Score + 
                     
                     Task_Difficulty + MW + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 
                                 
                                   dummy(MW) + 1 || ResponseId) + 
                               
                                 dummy(MW) + (1 || Stimulus),
                               
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_4.3)
```

### Comparing model_3.3 to model_4.3

```{r}
anova(model_3.3, model_4.3)
```

### Random effects of participants on intercepts and task difficulty and mind wandering slope, and stimulus items on intercepts and mind wandering slope with covariances

```{r}
model_5.3 <- glmer(Acc ~ DOCS_Score + 
                     
                     Task_Difficulty + MW + DOCS_Score:Task_Difficulty +
                               
                               (dummy(Task_Difficulty) + 
                                 
                                   dummy(MW) + 1 | ResponseId) + 
                               
                                 dummy(MW) + (1 | Stimulus),
                               
                             family = "binomial", 
                             glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                             
                             data = dat)

summary(model_5.3)
```

### Comparing model_4.3 to model_5.3

```{r}
anova(model_4.3, model_5.3)
```

## Visualising the results